---
title: "OTU_analysis"
author: "Alejandro Rojas"
date: "April 5, 2018"
output:
  html_document: 
    keep_md: TRUE
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("/Volumes/GoogleDrive/My Drive/Duke_vilgalys/SOB_project/Australia_amplicon/SOB_amplicon")
```

## Import data

Libraries used during this analysis:
```{r libs, message=FALSE, warning=FALSE}
library(phyloseq)
library(dplyr)
library(tidyr)
library(tibble)
library(ggplot2)
library(vegan)
library(readr)
```

Creating the otu table:
```{r otu_table, warning=FALSE}
otu_table_SOB <- read_delim("Complete_data/otu_table.fungi.txt", 
    "\t", escape_double = FALSE, trim_ws = TRUE)

#Separating objects into OTU table and taxa file
SOB_otu <- otu_table_SOB[,1:319] %>% column_to_rownames(var="OTU_ID") %>% as.matrix()
Pine_taxa <- otu_table_SOB[,c(1,320)] %>%
  separate(col = taxonomy, 
           into = c("taxa",
                    "Phylum",
                    "Class","Order",
                    "Family","Genus",
                    "Species"), sep = ";") %>%
  separate(col = taxa,
           into = c("spp", "genbank", "unite", "reps","Kingdom"), sep = "\\|") %>%
  select(c("OTU_ID","Kingdom", "Phylum",
                    "Class","Order",
                    "Family","Genus",
                    "Species")) %>% 
  column_to_rownames(var = "OTU_ID")
```

Based on the google doc, a metadata file was generated to merge the information with the OTU table and the taxonomy file.  This is the metadata assoicated with samples collected in Australia, we need to work on some of the codes to clean the information.  There are __318 samples__ collected and included in the current study.

```{r sample_data}
SOB_metadata.0 <- read.delim("metadata.txt", sep = "\t") %>% column_to_rownames(var = "SampleID")
SOB_metadata <- sample_data(SOB_metadata.0)
head(SOB_metadata)
```


Creating the phyloseq object to do the downstream analyses.

```{r phyloseq_obj}
#Creating OTU table
SOB_otu <- otu_table(SOB_otu, taxa_are_rows = TRUE)

#Creating taxa object
SOB_taxa <- tax_table(as.matrix(Pine_taxa))

#Creating phyloseq object
SOB_data <- phyloseq(SOB_otu, SOB_taxa, SOB_metadata)
SOB_data

#Removing unidentified or other kingdoms
SOB_data <- subset_taxa(SOB_data, Kingdom=="k__Fungi")
```

# Exploratory analyses

First data exploration after generating the phyloseq object in R.  Here is a plot of the abundance of each OTU colored by phyla on each of the locations sampled (there are few categories messed up).

## Bar plots at Phylum level
```{r}
#Stacked plot
fungi.p <- subset_taxa(SOB_data, Phylum!="p__unidentified")

fungi.p <- tax_glom(fungi.p, taxrank="Phylum")

bar_phyla <- plot_bar(fungi.p, x ="site_name", fill = "Phylum") 
bar_phyla + geom_bar(stat = "identity", position = "stack") + 
  scale_fill_brewer(type = "div", palette = "Paired") + coord_flip() 
```

Breaking the plot by type of location sampled and the type of forest sampled (INV: Invasive, PL: Plantation, and UN: Native).

```{r}
fungi.p2 <- subset_samples(fungi.p, site_code != "null")
code_names <- c(INV = "Invasive",
                PL = "Plantation",
                UN = "Native")

bar_phyla_code <- plot_bar(fungi.p2, x = "site_name", fill = "Phylum") 
bar_phyla_code + geom_bar(stat = "identity", position = "stack") + 
  facet_grid(site_code ~ ., labeller = labeller(site_code = code_names)) +
  scale_fill_brewer(type = "div", palette = "Paired")
```

## Alpha diversity
```{r}
alpha_SOB <- plot_richness(subset_samples(SOB_data, site_code != "null"),
                           x="site_code", measures = c("Chao1", "Shannon", "InvSimpson"))
alpha_SOB + geom_boxplot() + scale_x_discrete(labels = c("INV" = "Invasive", 
                                                         "PL" = "Plantation",
                                                         "UN" = "Native")) +
  xlab("Site")
```



## Ordination plots

```{r}
#Palette
library(RColorBrewer)
pal <- colorRampPalette(brewer.pal(12, "Paired"))

#Ordination
SOB_data.1 <- subset_samples(SOB_data, site_code != "null")
ord.SOB <- ordinate(SOB_data.1, "NMDS", "bray")
plot_ordination(SOB_data.1, ord.SOB, shape = "site_code", color = "site_name") + 
  geom_point(size = 3) + scale_color_manual(values = pal(42))

#NMDS after removing null samples
ord.SOB.1 <- ordinate(SOB_data.1, "NMDS", "bray")
plot_ordination(SOB_data.1, ord.SOB.1, shape = "site_code", color = "site_name") + 
  geom_point(size = 3) + geom_polygon(aes(fill=site_name), alpha = 0.3) +
  scale_color_manual(values = pal(42)) +
  scale_fill_manual(values = pal(42)) +
  facet_wrap(~state)
```

## Heatmap

Focusing on the top 100 OTUs, this is the distribution of OTUs per site.

```{r, fig.height=15, fig.width=10}
top100OTU.names <- names(sort(taxa_sums(SOB_data), TRUE)[1:100])

top100OTU <- prune_taxa(top100OTU.names, SOB_data)
heat_100OTU <- plot_heatmap(top100OTU,sample.label = "site_code", sample.order = "site_code", 
                            taxa.label = "Genus", taxa.order = "Phylum",
                            low = "white", high = "green", na.value = "grey")
heat_100OTU +  xlab("site code") + ylab("OTU genera ordered by phyla") +
  theme(axis.text.x = element_text(size = 9), 
        axis.text.y = element_text(size = 6))
```


## Rank abundance
This is a rank abundance plot of the top 100 OTUs:

```{r message=FALSE, warning=FALSE, fig.height=15}
library(ampvis)
amp_rabund(top100OTU, tax.aggregate = "Genus", tax.add = "Phylum", tax.show = 200,
           plot.log = TRUE) + ylab("log10(Relative abundance)")
```

If we limit this rank abundance to members of the phylum Basidiomycota, this is the distribution of the __top 150 Basidio OTUs__:

```{r message=FALSE, warning=FALSE, fig.height=15}
library(ampvis)
fungi.basidio <- subset_taxa(SOB_data, Phylum=="p__Basidiomycota")
top150OTU.names <- names(sort(taxa_sums(fungi.basidio), TRUE)[1:150])
top150OTU <- prune_taxa(top150OTU.names, fungi.basidio)

p1.ht <- ampvis::amp_rabund(top150OTU, tax.aggregate = "Genus", tax.add = "Phylum", 
          plot.log = TRUE) 
p1.ht + ylab("log10(Relative abundance)")
```


## Metacoder analysis

Using top 500 OTUs and focusing only in the agaricomycetes, this is the distribution of the different taxonomic levels below the __class Agaricomycetes__.

```{r, message=FALSE, warning=FALSE, cache=TRUE}

library(metacoder)

top500OTU.names <- names(sort(taxa_sums(SOB_data), TRUE)[1:500])

top500OTU <- prune_taxa(top500OTU.names, SOB_data)

#Converting to metacoder
obj <- parse_phyloseq(top500OTU)


# Convert counts to proportions
obj$data$otu_table <- calc_obs_props(obj,
                                     dataset = "otu_table",
                                     cols = obj$data$sam_data$sample_ids)
# Calculate per-taxon proportions
obj$data$tax_table <- calc_taxon_abund(obj,
                                       dataset = "otu_table",
                                       cols = obj$data$sam_data$sample_ids)

#Compare treatments
obj$data$diff_table <- compare_groups(obj,
                                          dataset = "tax_table",
                                          cols = obj$data$sam_data$sample_ids,
                                          groups = obj$data$sam_data$site_code)
```



```{r, fig.height= 14, fig.width=14, warning = FALSE}
#Tree visual
set.seed(1)
Tree1 <- metacoder::heat_tree(taxa::filter_taxa(obj, taxon_names == "c__Agaricomycetes", subtaxa = TRUE),
          node_size = n_obs, 
          node_label = taxon_names,
          node_color = log2_median_ratio,
          node_color_range = c("#a6611a","#dfc27d","#bdbdbd","#80cdc1","#018571"), 
          node_color_trans = "linear",
          node_label_max = 120,
          node_color_interval = c(-1, 1),
          edge_color_interval = c(-1, 1),
          node_size_axis_label = "Number of OTUs",
          node_color_axis_label = "Log2 ratio median proportions",
          initial_layout = "reingold-tilford", layout = "davidson-harel")

Tree1
```

```{r data_save, warning=FALSE, echo=FALSE}
save(SOB_data, top100OTU, obj, Tree1, file = "SOB_files.rda")
```

